### 2016-01-30 advantages.tex
Initial check in, FSE format

\subsection{Key Novel Features of Localization}

One key point is that the localization information provided by
repairing mutants is different in kind from that typically provided by
fault localization, particularly statistically-driven localizations.
Namely, the relationship between a mutant and the underlying fault is
\emph{always} causal.  There are not ``accidental'' repairs; at
minimum, a mutant modifies program behavior such that the faulty code
no longer causes a fault, which it would have under other
circumstances.  This means that unlike traditional localizations,
where reading ``incorrect'' suggestions (e.g., ones that are not an
actual faulty location) is often of no value, every mutant that
repairs a program provides some information about the fault, however
obscure.  Some repairs, of course, are of little interest because they repair too much.

A second point is that unlike the use of program modifications in
automatic patch generation, our technique is not defeated by weak
oracles.  Of course, if an oracle is too weak to detect failures at
all, it cannot be applied, but given the ability to detect failures,
it is not critical that a suite be able to check correctness.  Our
repairing mutants are not meant to serve as actual program patches,
but 1) as features to drive a fuzzer taming algorithm and 2) as error
explanations and fault localization to ease the debugging of complex
compiler faults.