### 2016-02-20 abstract.tex
abstract intro changes

Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.  This paper proposes a novel metric for measuring distance between executions, and specializes it to the most commonly needed case, failing test cases.  The metric is based on a causal concept of distance: executions are similar to the degree that changes in the program itself, introduced using mutation operators, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases (for the original program) become successful due to the same mutant, they are more likely to be due to the same fault.  This metric both improves an existing method for identifying test cases failing due to unique faults and enables novel fault localization approaches.  We evaluate our methods using more than 50 faults and 2,500 test cases for two complex, widely-used, real-world compilers, and demonstrate improvements over state-of-the-art methods.