### 2016-03-01 conclusions.tex
ready submit

\section{Conclusions and Future Work}
\label{conc}

When are two failing test cases due to the same fault?  This paper
demonstrates that
``when they are repaired by the same program mutants'' is a useful answer to
this question.  The problem of causality in executions is hard, and
has long been tied to a notion of similarity of
executions \cite{NearNeighbor,GroceError}.  By using program changes
(produced by mutants) as
causes to determine distance, it is possible to improve on efforts
to identify the set of faults in a large set of redundant failures
\cite{PLDI13} and discover the source code locations of faults
\cite{MUSE,multilingual}.

As future work, we plan to apply the mutation-based metric to other
fuzzer taming and fault localization problems.  The highly varying
results for even the best fault localization methods in our
experiments demonstrate that further advances are required before
compiler debugging can consistently be made less onerous through automated
assistance.  One idea is to combine the metric approach and the ideas
in MUSE and MUSEUM, in some way, to exploit passing tests and perhaps
additional coverage information.

%We also plan to explore other applications of mutant-based metrics.
Many software engineering techniques rely on measuring distance
between program executions \cite{BallConcept,NearNeighbor}.  Not all
of these concern only failing
executions.  For example, Zhang et al. \cite{issta14} show that FPF based on a simple
Hamming distance over branches covered can significantly improve the
effectiveness of seeded symbolic execution
\cite{Zesti,PersonSeed,BugRedux}.  Mutation response is likely too
expensive to use for this purpose without modification, but sampling
mutants to refine distance when coarser metrics plateau
may be practical.  Mutants may also be useful in determining
execution peers \cite{Sumner2011}, for example to help operators of
spacecraft find similar behaviors to telemetry downlinks in testbed
history \cite{KlausRajeev,scriptstospecs}.