### 2016-03-01 metrics.tex
ready submit

\section{A Mutant-Based Distance Metric}

For every failing test case, its response to mutants defines a bit-vector, with
length equal to the number of mutants that repaired any test case,
where a 1 bit indicates that the mutant in question repairs this test
case.  These cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   Hamming distance for the first pair of tests is twice as
large, despite the fact we have no evidence $t_1'$ and
$t_2'$ are due to the same fault, in contrast to strong evidence for
$t_1$ and $t_2$.  A Jaccard distance, also used (as similarity) in some fault
localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

This metric, the portion of mismatches over bits that are 1 in either
vector, matches our intuitions about the above comparisons.  It
essentially formalizes the intuition of the onion-ring model, where
more matching repairs indicates a much higher probability two failures
are due to the same fault, even if not all repairs match.

Given such a metric, FPF \cite{Gonzalez} fuzzer taming \cite{PLDI13}
begins by ranking as first an arbitrary test case.   FPF next ranks the test
case, of all unranked tests, that has the largest minimum distance
from all ranked test cases --- the test case most dissimilar by \emph{closest}
already ranked test --- until all tests are ranked.  FPF is boundedly close
to optimal under certain assumptions \cite{Gonzalez}.

More complex metrics are possible, also combining information such as
coverage and language features or test case tokens \cite{PLDI13}.
However, mixtures of metrics (combining different features) other than
Levenshteins \cite{lev} over test case + output never performed well
compared to more focused metrics in earlier work \cite{PLDI13}.
Therefore, in order to evaluate the contribution of our repair-based
metric we compare it, alone, to the previous best metrics.

In Section \ref{conc} we briefly discuss possible uses of $d$
requiring comparison of both failing and passing test cases. The only
change required is to use three-valued vectors, where 1 indicates a failing property succeeds and -1 indicates a succeeding property fails.