### 2016-05-03 response.txt
ready for fse submit

We thank the reviewers for very helpful comments.

R1:

...significant over PLDI 2013...

Using our samplings and top-N PLDI 13 results (non-ideal method, hence not in paper), results _are_ _statistically_ significant (know not what you meant, but worth noting).

Q: ...correlation between... test cases and... invariants?

Even reduced test cases are hard to connect to invariants, tend to execute ~10-50K+ LOC (https://github.com/agroce/mutants16/tree/master/tests).  Probably is a correlation, but it's hard to spot for compiler tests.  We'll definitely investigate further!

Q: ...optimizations... majority of bugs...? ...non-compiler benchmarks?

~90% of GCC faults detected by changing from -O0 to -O3, so seems to be the rule for GCC.  For SpiderMonkey, "optimization" hard to separate from odd JavaScript semantics, or attempts to speed up SpiderMonkey in general (not always traditional optimization).  Results of Qi et al. suggest many programs _can_ be "patched" by turning off a feature; MUSEUM results also suggest generalization plausible.

3) ...baseline probability. What does this mean?

(Baseline) B = chance _any_ two failing tests are due to same fault.  Knowing two failures are "fixed" by turning off same optimization indicates probability they are same fault is > B.  We quantify (in 4.1, 4.2) this claim (though not for optimizations alone).

R2:

Q: ...center... repairs many others? Will that fault be discovered?

Really depends on the overall structure (what repairs which other tests).

Q: ...faults not discovered in the experiment...?

No obvious patterns, but differences would be complex properties of C/JavaScript and GCC/SpiderMonkey.  We'll investigate further!

Q: ...sampling does alter the results... undesirable happening in terms of the onion ring model.

Sampling hurts localization, maybe ok for taming (we're investigating).  Our guess: localization is bad when nothing in the middle of the onion ring is left (could be small set), which hurts taming less.

We're very interested in alternatives to general mutation, thanks for ideas!

Will provide runtimes (JavaScript ok, GCC bad, but we ran/built inefficiently).

R3:

TACAS04 is completely different metric, used very differently, we believe.

Q: ...what kind of changes... ...insert System.exit(0)...?

Anything such that the test oracle passes is allowed.  For a very weak oracle, you could get away with this.  Due to onion ring, letting some "silly" changes make tests pass is probably harmless, so long as other more interesting mutants are also used.

Q: ...don’t quite understand... evaluation results...

Figure 2 shows _only_ fuzzer taming results, no localization; Tables show _only_  localization results.  These are _independent_ goals, using same metric.  FPF column only in Tables to show that how well a fault localizes and how well it is identified seem (surprisingly) unrelated.

Q: ...role of passing tests... "lower" than... failing tests?

We believe so: probably for fuzzer taming, possibly for localization (MUSEUM and other localizations agree).  Faulty code often executes in many passing tests; determining conditions for failure is difficult; thus more information _about faults_ in failures, in some sense: e.g., single failing test can show a fault exists, which all passing tests cannot!  We use passing tests to prune mutants.

Q: ...how to interpret... Figure 3...

992 (commented "ADDED") is part of the code in the patch, as is the deleted statement (line #s in patch would help).  We'll clarify this much better, sorry!  3/6 is indeed ranked change.

Q:  ...confirm that these 8 faults are correct...?

For these 8 (and 16 GCC faults), we had a backpatch, based on commit data and comments, and tested it.  The other faults had such complex change history this was not feasible, making localization (not taming) evaluation impossible.  Backpatch sets were produced by colleagues, not us, before we began this work, to avoid bias.