### 2019-01-20 journal/example.tex
mutants16 paper

\section{A Simple Example}

To better understand the causal approach taken in this paper, it is useful to consider a simplified, abstract example, and understand how the approach proposed would apply to this example.