### 2019-01-23 journal/example.tex

\section{A Simple Example}
\label{sec:example}

To better understand the causal approach taken in this paper, it is useful to consider a simplified, abstract example, and understand how the approach proposed would apply to this example.

\begin{figure*}
\begin{code}
 01  void compile(program: feature list, optimizeLevel: int) \{
 02    internalRepresentation = initInternalRep(program);
 03    if optimizeLevel > 0: \{
 04       if feature1 in program:
 05          if feature2 in program:
 06             corrupt(internalRepresentation);
 07       if optimizeLevel > 1: \{
 08          if feature3 in program:
 09             if feature4 not in program:
 10                corrupt(internalRepresentation);
 11          if feature5 in program:
 12             if feature6 in program:
 13                corrupt(internalRepresentation);
 14       \}
 15    \}
 16    /* Will produce wrong code if structure has been corrupted */
 17    generateCode(internalRepresentation);
\end{code}
\caption{A toy ``optimizing compiler'' with multiple bugs}
\label{fig:example}
\end{figure*}

\begin{figure}
\begin{code}
T1: (1, 2)
T2: (3)
T3: (5, 6)
\end{code}
\caption{Failing tests for the toy compiler.}
\label{fig:toybugs}
\end{figure}

Consider the pseudo-code ``compiler'' in Figure \ref{fig:example}.  We can imagine that there is a large amount of code (not shown) in the {\tt initInternalRep} and {\tt generateCode} functions.  Assume, however, that the only bugs in the compiler are the {\tt corrupt} calls.  Traditional Spectrum-Based Fault Localization (SBFL) techniques can be easily ``confused'' in multiple-fault settings, and by cases such as this, where there may be strong correlations between code covered by the (fault-free) {\tt initInternalRep} and {\tt generateCode} functions and the features contained in programs, making lines of code far from the fault locations appear suspicous.

Assume that all of our test cases run the compiler with {\tt
  optimizeLevel} of 2, so as to potentially apply all optimizations.
A popular way to fuzz a compiler is to compare program behavior when
code is compiled with all optimizations with program behavior when
code is compiled with no optimizations, so this reflects real-world compiler testing.  A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.

Further, let us only consider two kinds of mutants:
condition-negation and statement-deletion mutants.  Obviously,
negating the condition on line 3 will make all of our failing test
cases pass, and negating the condition on line 7 will make test cases
2 and 3 pass, but not test case 1.  Negating the conditions on lines
4-5, 8-9, and 11-12 will repair our test cases, but these changes are
also likely to make some passing test case fail (they represent
preconditions for compiler optimizations), so these mutants will not
be used in our analysis.  Finally, the only statements that can be deleted without breaking the code are deletions of the {\tt corrupt} calls on lines 6, 10, and 13.  We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.  Each test can then be associated with a bitvector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.  The vectors for our tests, then, are:

\begin{code}
T1: 11000
T2: 10110
T3: 10101
\end{code}

T1 thus has a distance of 0.75 from T2 and T3.  T2 has
a distance of only 0.5 from T3, however.

\subsection{Fuzzer Taming for the Toy Compiler}

If the Furthest-Point-First
algorithm starts from T1 it will arbitrarily select either T2
or T3 next, then the other test.  If it starts with T2 or T3, it
will select T1 next, then the other of the two closer tests.  Of
course, taming a set of three failing tests is not difficult!
However, if we imagine a much larger set of failing tests, which
contain irrelevant features that cannot be removed by delta-debugging
without causing the programs to become syntactically invalid, so long
as the basic structure of the cause for failure remains the same, the
the FPF algorithm will still select tests with the same bitvectors as
T1, T2, and T3.  Concretely, if we add these tests (shown with associated bitvectors):

\begin{code}
T4: (1, 2, 7)    11000
T5: (3, 6)       10110
T6: (1, 5, 6)    10101
T7: (1, 2, 5, 6) 10000
\end{code}

\noindent then FPF from T1 will proceed to a maximally distant
test, such as T5, at distance 0.75.  The next ranked test could not have
been T4 (distance 0.0) or T7 (distance 0.5), but could have
been T2, T3, or T6.  After T1 and T5, the next test in the ranking will be the test whose nearest
neighbor among already chosen tests is furthest away.  The distances
to the two tests already chosen are:

{\scriptsize
\begin{code}
                         11000  10110
T2: (3)          10110   0.75   0.00
T3: (5, 6)       10101   0.75   0.50
T4: (1, 2, 7)    11000   0.00   0.75
T6: (1, 5, 6)    10101   0.75   0.50
T7: (1, 2, 5, 6) 10000   0.50   0.67
\end{code}
}

\noindent The FPF algorithm could choose either T3, T6, or T7 all of
which have a nearest neighbor at distance 0.5.  Let
us assume it chooses T6.  At this point, the three distinct bugs
are all ranked.  However, there is one interesting pattern remaining.
The distances to tests are now:

{\scriptsize
\begin{code}
                         11000  10110  10101
T2: (3)          10110   0.75   0.00   0.50
T3: (5, 6)       10101   0.75   0.50   0.00
T4: (1, 2, 7)    11000   0.00   0.75   0.75
T7: (1, 2, 5, 6) 10000   0.50   0.67   0.67
\end{code}
}

\noindent The algorithm is forced to choose T7, the only remaining
test that does not have the same bitvector as another test.  T7 is
worth ranking, as it shows that a program can be written that will
result in wrong code unless both bugs it triggers are fixed.

\subsection{Fault Localization for the Toy Compiler}

Now let us consider the {\bf Repair} fault localization algorithm
applied to this example.  If we are trying to fix the problem
exhibited by T2, what will the system suggest to us as the
location of the fault?  Recall that 1) in {\bf Repair}, a localization
is the location (and content) of a repairing mutant.  T2 is repaired
by negating the condition on line 3, negating the condition on line 7,
or deleting the statement on line 10.  Which of these is best?  {\bf
  Repair} distinguishes between repairing mutants by ordering them by
their distance to the most-distant (by bitvector) test also repaired
by that mutant.  The only other tests repaired by deletion of line 10
are the other instances of the same bug, all at distance 0.  Of the
other two repairs, negating the condition on line 3 also repairs T1,
at distance 0.75, while negating the condition on line 2 repairs
nothing farther away than T3 at 0.5.
Therefore, the localization for T2 is (in rank order):

{\scriptsize {\tt 1: DELETE  10                corrupt(internalRepresentation);}}

{\scriptsize {\tt 2: NEGATE  07       if optimizeLevel > 1: \{}}

{\scriptsize {\tt 3: NEGATE  03    if optimizeLevel > 0: \{}}

\subsection{The Onion Ring}

In this simple toy example, the onion ring will only have a few
layers, one for each possible distance from the test at the center.
For example, if we take T2 as the center, then the layers will be:

\begin{code}
 0.00:  T2 T5
 0.50:  T3 T6
 0.67:  T7
 0.75:  T1 T4
\end{code}