### 2019-10-01 journal/metrics.tex
small fixes' git push

\section{A Mutant-Based Distance Metric}

For every failing test case, the response (fixing or not fixing the
failure) to all mutants defines a bitvector, with
length equal to the number of mutants that repaired \emph{any} test
case (or all mutants; it does not matter for our proposed metric),
where a 1 bit indicates that the mutant in question repairs this test
case\footnote{By \emph{repairing} a failing test we here mean that a mutant
causes the test to pass, \emph{and} the mutant does not cause
other, previously passing tests, to fail.}.  These bitvectors cannot be effectively compared with Hamming distance. Consider $t_1$ and $t_2$, both repaired
by 100 mutants.  If these two test cases share 90 common repairing
mutants, and disagree for 10, they have a Hamming
distance of 20.  This makes them ``less similar'' than two test cases,
$t_1'$ and $t_2'$, where $t_1'$ is repaired by 5 mutants, $t_2'$ is
repaired by 5 mutants, and \emph{none of those mutants overlap.}   A Jaccard distance, also used (as similarity) in some fault
localization algorithms \cite{Pinpoint}, is a better fit.  For bitvectors $u$ and $v$ of length $n$:
%\vspace{-0.08in}
\[d(u,v) = 
\begin{cases}
\frac{\sum_{i=0}^{n-1} 1\ \text{if}\ u_i \neq v_i\ \text{else}\ 0}{\sum_{i=0}^{n-1} 1
\  \text{if}\ u_i \neq 0 \vee v_i \neq 0\ \text{else}\ 0} & \exists i:u_i\neq 0 \vee v_i\neq 0\\
0 & otherwise
\end{cases}
\]

\noindent In other words, the Jaccard distance is $\frac{\mathit{bitcount}(u
\veebar v)}{\mathit{bitcount}(u \vee v)}$ (the number of 1 bits in the
XOR over the number of 1 bits in the OR for $u$ and $v$); that is, the
distance is the portion of mismatches over repairing mutants for
either test.  This metric agrees with our intuitions about the above comparisons.  It
essentially formalizes the intuition of the onion-ring model, where
more matching repairs indicates a much higher probability two failures
are due to the same fault, even if not all repairs match.

Given such a metric, FPF \cite{Gonzalez} fuzzer taming \cite{PLDI13}
begins by ranking as first an arbitrary test case.   FPF next ranks the test
case, of all unranked tests, that has the largest minimum distance
from all ranked test cases --- the test case most dissimilar by \emph{closest}
already ranked test --- until all tests are ranked.  FPF is boundedly close
to optimal under certain assumptions \cite{Gonzalez}.

More complex metrics are possible, also combining information such as
coverage and language features or test case tokens \cite{PLDI13}.
However, mixtures of metrics (combining different features) other than
Levenshteins \cite{lev} over test case + output never performed well
compared to more focused metrics in earlier work \cite{PLDI13}.
Therefore, in order to evaluate the contribution of our repair-based
metric we compare it, alone, to the previous best metrics.

In Section \ref{conc} we discuss uses of $d$
requiring comparison of both failing and passing test cases:  in this 
case, we would use three-valued vectors, with a bit for each \emph{property}, where 1 indicates a 
failing property succeeds and -1 that a succeeding property 
fails.  For such an application, we could remove the requirement that
a repair not be a killed mutant, since the passing test vectors would encode
that information